%Schemaless schema 
\chapter{Data Model}

\input{er.tex}
\input{querying_requirements.tex}

\section{Database schema}
JAQPOT Quattro will make use of a document database, 
which is to a good extent schema-less. Here we discuss about 
the basic documents of JAQPOT Quattro and the indexes we need to 
introduce so as to maximize the performance of the 
search, insert and update operations defined previously.

The overall data schema refers to three databases:

\begin{enumerate}
 \item  A document database, in our case mongodb, that stores data in the form of documents. Most of JAQPOT’s resources are stored in this database,
\item  A relational database system, here mysql, that stores certain relations that have to do with datasets
\item  An ElasticSearch DB system that stores keyword-related data to facilitate the exploration of our domain.
\end{enumerate}

\subsection{Task and ErrorReport}

\textit{Example 1.} A completed task (note: duration is in milliseconds).

\begin{lstlisting}[language=json]
{ 
    "status":"COMPLETED", 
    "http_status":200, 
    "result": "http://opentox.ntua.gr:8080/model/234", 
    "duration":452, 
    "progress":[ 
        "data loaded", 
        "new resources have been created", 
        "now training the model", 
        "model training completed", 
        "result: OK" 
        ], 
    "creator":"guest@opentox.ntua.gr", 
    "time_created":"2015-12-01 14:04:45 CET" 
} 
\end{lstlisting}

\noindent \textit{Example 2.} A failed task with an error report in it:
\begin{lstlisting}[language=json]
{ 
    "status":"ERROR", 
    "http_status":400, 
    "duration":200, 
    "progress":[ 
        "data loaded", 
        "new resources have been created", 
        "now training the model", 
        "failed!" 
        ], 
    "creator":"guest@opentox.ntua.gr", 
    "time_created":"2015-12-01 14:04:45 CET", 
    "error_report":{ 
        "error_code":"XR243", 
        "actor":"client with IP 10.8.0.4", 
        "message":"Malformed input data", 
        "details":"Provide some details here to assist debugging", 
        "trace": { 
            "comment":"some other error report goes here if necessary" 
        } 
    } 
} 
\end{lstlisting}

\noindent \textit{Example 3.} A running task:
\begin{lstlisting}[language=json]
{ 
    "status":"RUNNING", 
    "percentage":57, 
    "http_status":202, 
    "progress":[ 
        "data loaded", 
        "new resources have been created", 
        "now training the model" 
        ], 
    "creator":"guest@opentox.ntua.gr", 
    "time_created":"2015-12-01 14:04:45 CET" 
} 
\end{lstlisting}

Notice that we do not assign IDs to the above documents since 
mongodb will create automatically a unique ID which is always 
denoted by \texttt{\_id}.

Indexes:
1. creator
2. status
3. (creator, status)

\begin{lstlisting}[language=json]
db.tasks.ensureIndex({"creator":1});
db.tasks.ensureIndex({"status":1});
db.tasks.ensureIndex({"creator":1, "status":1});
\end{lstlisting}

One can update the metadata of a task. 
See this example:
\begin{lstlisting}[language=json]
db.tasks.update( 
  { _id: ObjectId("54b454f8feee5a13a43f34f3") },  
  { $addToSet:  
     { "progress": "moved to step 4" } 
  } 
); 
\end{lstlisting}

All other update operations can be performed in a very standard way.

It is possible to fetch the data of a document from the DB without the data in a given field. For example, we can fetch a task without the progress information (here we assume we know the ID of the task):

\begin{lstlisting}[language=json]
db.tasks.find(
  {_id: ObjectId("54b454f8feee5a13a43f34f3")}, 
  {progress:0}
).pretty();
\end{lstlisting}

in some cases we may need to check only the status of a task (provided that we know its ID):
\begin{lstlisting}[language=json]
db.tasks.find(
  {_id: ObjectId("54b454f8feee5a13a43f34f3")}, 
  {status:1}
).pretty();
\end{lstlisting}

Remarks:
1. Tasks should be implemented as TTL (time-to-live) collections so that old tasks are garbage-collected.